Search-Based Test Data Generation for JavaScript Functions that Interact with the DOM \thanks{Identify applicable funding agency here. If none, delete this.}

### 2018-08-12 doc/issre2018/conclusions.tex
Working on camera-ready. Fixed all comments by Doatse.

In this paper we have introduced a novel JS unit testing framework, called \emph{JEDI}, which uses search-based techniques for generating test data. Notably, our testing framework is able to generate an arbitrary DOM input, which is syntactically valid and compliant with the recent HTML5 specification. We propose a novel test generation algorithm, called ``genetic with restart'', combining the benefits of systematic fitness improvement with the genetic search and the ability of escaping plateau by restarting search with a new target. Based on the case studies found in two related works~\emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011}, we conduced an empirical validation of our framework and performed a significance study of the results. The validation has shown the effectiveness of the framework in covering target branches with a reasonable efficiency.

>>>>

In this paper we have introduced a novel JS unit testing framework, called \emph{JEDI}, which uses search-based techniques for generating test data. Notably, our testing framework is able to generate an arbitrary DOM input, which is syntactically valid and compliant with the recent HTML5 specification. We propose a novel test generation algorithm, called ``genetic with restart'', combining the benefits of systematic fitness improvement with the genetic search and the ability of escaping plateaus by restarting search with a new target. Based on the case studies found in two related works~\emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011}, we conduced an empirical validation of our framework and performed a significance study of the results. The validation has shown the effectiveness of the framework in covering target branches with a reasonable efficiency.

### 2018-08-12 doc/issre2018/evaluation.tex
Working on camera-ready. Fixed all comments by Doatse.

In order to evaluate the feasibility of our test framework, we conducted an empirical study on a set of subjects (see Section~\ref{sub.sec.case.studies}). All experiments were submitted as a batch mode job on a cluster machine with 32 HT Cores and 256 GB RAM running under Scientific Linux 7.3. We have created three different instances of the JS testing framework --- each one is based on a slight variation of the underlying genetic algorithm. The first version, which plays the role of our baseline, is \emph{Random} ($\Random$). Instead of applying crossover and mutation to evolve the population, this algorithm regenerates the new population randomly from scratch. It is important to note  that the random version also benefits from both preliminary static analysis and dynamic constant propagation as part of the fitness evaluation. The second version is \emph{Genetic}~($\Genetic$) --- it literally follows all of the steps prescribed by the genetic framework in Figure~\ref{fig.framework.architect}. The last version is \emph{Genetic with Restart} ($\RGenetic$). It is based on the genetic Algorithm~\ref{alg.gen} where the search is restarted after the archives are converged within a certain history window. On the one hand, the smaller window size forces the GA to converge faster and quickly discovers stagnation. On the other hand, the small window can lead to premature archive convergence, which prevents the right genetic combination from being triggered. We have experimented with two different windows of size 5 and 10. But since the latter one ($\RGenetic$) showed the best performance over all, we chose to report only that case in the paper. Table~\ref{tbl.gen.config} summarizes the configuration parameters for all three versions of the testing framework. The initial setup for genetic algorithm is similar to other frameworks, e.g. \emph{EvoSuite}~\cite{fraser2011evosuite}.

@@

Being one of the most popular programming languages for web development, there are ample open source JS projects on the Web. JS applications are written with the help of countless numbers of specialized libraries, and greatly vary in size, complexity and coding style. Moreover, some popular JS web-frameworks, such as jQuery and React~\cite{todomvc}, implement their own API for DOM manipulation. Before selection, the source code of a potential case-study has to be scrutinized which is a time consuming and labor intensive process. Therefore, we decided to primarily focus on the evaluation of the JS subjects already established in the related research such as \emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011, tajsbenchmarks}. Both of these works address the problem of testing of JS application with the reference to the browser DOM. We added additional case-study \emph{mathjs}~\cite{mathjs} to increase the presence of primitive (browserless) JS code. For each case-study application, we identified several JS functions that are suitable for our testing framework. In the end, our evaluation set included both primitive and DOM related JS functions.

@@

To study the effectiveness and efficiency of our test generation framework, we exercised each algorithm ($\Random$, $\Genetic$ and $\RGenetic$) against all branches of the functions in Table~\ref{tbl.case.studies}. Due to the probabilistic nature of the test generation process, the experiments have been repeated 50 times to achieve statistical significance. Table~\ref{tbl.stats} summarizes the evaluation results. For each algorithm, we computed the following \emph{medians} per branch: the number of iterations~($i$), time in seconds~($t$), and their ratio~($t/i$). In the table, these values are grouped into three columns $m_{\Random}$, $m_{\Genetic}$ and $m_{\RGenetic}$ one for each algorithm. The next group of columns $t_{\Random}/t_{\Genetic}$, $t_{\Random}/t_{\RGenetic}$ and $t_{\Genetic}/t_{\RGenetic}$ presents the relative performance of each pair of the testing algorithm $(\Random,\Genetic)$, $(\Random,\RGenetic)$, and $(\Genetic,\RGenetic)$ respectively. To the right of the function name, we also present accumulated statistics for the function as a whole\footnote{because of the missing branches, the function's summary may not be equal to the sum of the branches, e.g. see the \emph{toggleInfo} function}

@@

\textbf{RQ1 (effectiveness)}: To answer this question, we measure the branch coverage achieved by each test generation algorithm. Given a fixed limit on the number of test iterations, 200 in our case, a search algorithm can not guarantee finding a solution. The underlined values in Table~\ref{tbl.stats} like $\underline{200}$ indicate branches which were not covered by the respective algorithms. The main reason for the random generation~($\Random$) to fail is the large size of the search space and the lack of any feedback in the generation process, e.g. branch $(10,11)$ of the \emph{isGameFinished} function. In the case of the genetic approach~($\Genetic$), it  may get stuck at a plateau because the fitness function does not provide sufficient guidance, whereas $\Random$ could still escape the plateau by sheer luck. We encountered three such cases in our study: branches $(10,11)$, $(47,48)$ and $(45,45)$ of the function \emph{doPaddlePower}, \emph{initBricks} and \emph{do_draw}, respectively. They motivated us to introduce the genetic algorithm with restart ($\RGenetic$) which combines the strengths of both $\Random$ and $\Genetic$. It leverages the genetic search while the fitness continues to improve, and restarts the search with a new target, when the fitness suboptimally converges. As our evaluation has shown, $\RGenetic$ not only managed to outperform $\Random$ and cover the three problematic branches above, but it has generally improved the results of $\Genetic$, e.g. branch $(69,71)$ of the \emph{initBricks} function.\\
\framebox{
  \parbox[t][1.1cm]{0.95\linewidth}{
    Across all subjects, the $\RGenetic$ algorithm achieved 100\% branch coverage, with $\Genetic$ in the second place with 95\% coverage, and, finally, $\Random$ with  63\% coverage.
  }
}

@@

Another question related to the efficiency is the computational cost of one iteration of each algorithm. It is useful to know in order to choose the generation limit. The cost, shown in column $t/i$ in Table~\ref{tbl.stats}, is calculated as the relation between the execution time and the number of generations, i.e. it measures the speed of one iteration. Figure~\ref{fig.get.cost} presents the cost comparison for all three algorithms in our study. Among the factors impacting the iteration cost are the population size (50 candidates in our case) and the fitness evaluation of one candidate. The length of an execution trace correlates with the fitness evaluation. So, an exceptional program termination leads to short traces, whereas loops tend to produce longer traces. For example, the \emph{do_draw} function has long traces because it only operates with the primitive types thus never rising exceptions, and at the same time, it consists of 40 LOC. In our evaluation, the iteration cost varied between 0.25 and 3.59 seconds.\\
\framebox{
  \parbox[t][0.7cm]{0.95\linewidth}{
    On average, one algorithm iteration took one second for $\Random$ and $\RGenetic$, $\Genetic$ performed somewhat worse at 1.4 seconds.
  }
}

\textbf{RQ3 (significance):} Our testing framework is randomized by nature, therefore we need to conduct a statistical analysis to assess the \emph{significance} and \emph{effect size} of our results. This analysis is based on 50 executions produced by each test algorithm $\Random$, $\Genetic$ and $\RGenetic$. Following recommendations on the assessment of randomized algorithms~\cite{arcuri2011practical}, we use the non-parametric Mann-Whitney U-test~\cite{mann1947test} and the Vargha-Delaney $\hat{A}_{12}$ statistics~\cite{vargha2000critique} for measuring statistical significance ($\alpha=0.05$) and effect size, respectively. In Table~\ref{tbl.stats} the three columns to the right show the $\hat{A}_{12}$ values for all algorithm combinations, where this value is significant and '-' otherwise. Thus, the higher the value in the $_{\Genetic}^{\Random}$ column, indicates that $\Random$ is slower comparison with $\Genetic$. Depending on the actual $\hat{A}_{12}$ value, the effect size can be small (0.56), medium (0.64), and large (0.71).\\
\framebox{
  \parbox[t][0.65cm]{0.95\linewidth}{
    Both $\Genetic$ and $\RGenetic$ largely outperform $\Random$ in 50\% cases generally across the board and 67\% on the difficult functions.
  }
}

@@

\textbf{Threats to Validity:}
The \emph{construct validity} threat comes from the choice of measurements: branch coverage and execution time. Although, the branch coverage is a common criterion used in practice, the actual fault finding capability is a stronger indication of testing adequacy. Reaching a target branch is only part of the problem: we also need an oracle to discriminate the program's output. In this paper, we strike for so called ``natural'' oracles leading to exceptions. Using a stronger oracle for validation is future work. Another practical aspect of the test generation is the test case size, since it can both complicate the readability and increase the execution time. To mitigate this issue we configured the GA to start evolution with the candidates of small sizes, which helps to prevent on unnecessary data explosion. Additional measure that could be applied is a minimization procedure, aka \emph{hierarchical delta debugging}~\cite{misherghi2006hdd} and its variations~\cite{herfert2017automatically}, to the GA result.

>>>>

In order to evaluate the feasibility of our test framework, we conducted an empirical study on a set of subjects (see Section~\ref{sub.sec.case.studies}). All experiments were submitted as a batch mode job on a cluster machine with 32 HT Cores and 256 GB RAM running under Scientific Linux 7.3. We have created three different instances of the JS testing framework --- each one is based on a slight variation of the underlying genetic algorithm. The first version, which plays the role of our baseline, is \emph{Random} ($\Random$). Instead of applying crossover and mutation to evolve the population, this algorithm regenerates the new population randomly from scratch. It is important to note  that the random version also benefits from both preliminary static analysis and dynamic constant propagation as part of the fitness evaluation. The second version is \emph{Genetic}~($\Genetic$) --- it literally follows all of the steps prescribed by the genetic framework in Figure~\ref{fig.framework.architect}. The last version is \emph{Genetic with Restart} ($\RGenetic$). It is based on the genetic Algorithm~\ref{alg.gen} where the search is restarted after the archives are converged within a certain history window. On the one hand, the smaller window size forces the GA to converge faster and quickly discovers stagnation. On the other hand, the small window can lead to premature archive convergence, which prevents the right genetic combination from being triggered. We have experimented with the windows of size 5 and 10. But since the latter one ($\RGenetic$) showed the best performance over all, we chose to report only that case here. Table~\ref{tbl.gen.config} summarizes the configuration parameters for all three versions of the testing framework. The initial setup for genetic algorithm is similar to other frameworks, e.g. \emph{EvoSuite}~\cite{fraser2011evosuite}.

@@

Being one of the most popular programming languages for web development, there are ample open source JS projects on the Web. JS applications are written with the help of countless specialized libraries, and greatly vary in size, complexity and coding style. Moreover, some popular JS web-frameworks, such as jQuery and React~\cite{todomvc}, implement their own API for DOM manipulation. Before selection, the source code of a potential case-study has to be scrutinized which is a time consuming and labor intensive process. Therefore, we decided to primarily focus on the evaluation of the JS subjects already established in the related research such as \emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011, tajsbenchmarks}. Both of these works address the problem of testing of JS application with the reference to the browser DOM. We added additional case-study \emph{mathjs}~\cite{mathjs} to increase the presence of primitive (browserless) JS code. For each case-study application, we identified several JS functions that are suitable for our testing framework. In the end, our evaluation set included both primitive and DOM related JS functions.

@@

To study the effectiveness and efficiency of our test generation framework, we exercised each algorithm ($\Random$, $\Genetic$ and $\RGenetic$) against all branches of the functions in Table~\ref{tbl.case.studies}. Due to the probabilistic nature of the test generation process, the experiments have been repeated 50 times to achieve statistical significance~\cite{arcuri2011practical}. Table~\ref{tbl.stats} summarizes the evaluation results. For each algorithm, we computed the following \emph{medians} per branch: the number of iterations~($i$), time in seconds~($t$), and their ratio~($t/i$). In the table, these values are grouped into three columns $m_{\Random}$, $m_{\Genetic}$ and $m_{\RGenetic}$ one for each algorithm. The next group of columns $t_{\Random}/t_{\Genetic}$, $t_{\Random}/t_{\RGenetic}$ and $t_{\Genetic}/t_{\RGenetic}$ presents the relative performance of each pair of the testing algorithm $(\Random,\Genetic)$, $(\Random,\RGenetic)$, and $(\Genetic,\RGenetic)$ respectively. To the right of the function name, we also present accumulated statistics for the function as a whole\footnote{because of the missing branches, the function's summary may not be equal to the sum of the branches, e.g. see the \emph{toggleInfo} function}

@@

\textbf{RQ1 (effectiveness)}: To answer this question, we measure the branch coverage achieved by each test generation algorithm. Given a fixed limit on the number of test iterations, 200 in our case, a search algorithm cannot guarantee finding a solution. The underlined values in Table~\ref{tbl.stats} like $\underline{200}$ indicate branches which were not covered by the respective algorithms. The main reason for the random generation~($\Random$) to fail is the large size of the search space and the lack of any feedback in the generation process, e.g. branch $(10,11)$ of the \emph{isGameFinished} function. In the case of the genetic approach~($\Genetic$), it  may get stuck at a plateau because the fitness function does not provide sufficient guidance, whereas $\Random$ could still escape the plateau by sheer luck. We encountered three such cases in our study: branches $(10,11)$, $(47,48)$ and $(45,45)$ of the function \emph{doPaddlePower}, \emph{initBricks} and \emph{do_draw}, respectively. They motivated us to introduce the genetic algorithm with restart ($\RGenetic$) which combines the strengths of both $\Random$ and $\Genetic$. It leverages the genetic search while the fitness continues to improve, and restarts the search with a new target, when the fitness suboptimally converges. As our evaluation has shown, $\RGenetic$ not only managed to outperform $\Random$ and cover the three problematic branches above, but it has generally improved the results of $\Genetic$, e.g. branch $(69,71)$ of the \emph{initBricks} function.\\
\framebox{
  \parbox[t][1.1cm]{0.95\linewidth}{
    Across all subjects, the $\RGenetic$ algorithm achieved 100\% branch coverage, with $\Genetic$ in the second place with 95\% coverage, and, finally, $\Random$ with  63\% coverage.
  }
}

@@

Another question related to the efficiency is the computational cost of one iteration of each algorithm. It is useful to know in order to choose the generation limit. The cost, shown in column $t/i$ in Table~\ref{tbl.stats}, is calculated as the relation between the execution time and the number of generations, i.e. it measures the speed of one iteration. Figure~\ref{fig.get.cost} presents the cost comparison for all three algorithms in our study. Among the factors impacting the iteration cost are the population size (50 candidates in our case) and the fitness evaluation of one candidate. The length of an execution trace correlates with the fitness evaluation. So, an exceptional program termination leads to short traces, whereas loops tend to produce longer traces. For example, the \emph{do_draw} function has long traces because it only operates with the primitive types thus never rises exceptions, and at the same time, it consists of 40 LOC. In our evaluation, the iteration cost varied between 0.25 and 3.59 seconds.\\
\framebox{
  \parbox[t][0.7cm]{0.95\linewidth}{
    On average, one algorithm iteration took one second for $\Random$ and $\RGenetic$, $\Genetic$ performed somewhat worse at 1.4 seconds.
  }
}

\textbf{RQ3 (significance):} Our testing framework is randomized by nature, therefore we need to conduct a statistical analysis to assess the \emph{significance} and \emph{effect size} of our results. This analysis is based on 50 executions produced by each test algorithm $\Random$, $\Genetic$ and $\RGenetic$. Following recommendations on the assessment of randomized algorithms~\cite{arcuri2011practical}, we use the non-parametric Mann-Whitney U-test~\cite{mann1947test} and the Vargha-Delaney $\hat{A}_{12}$ statistics~\cite{vargha2000critique} for measuring statistical significance ($\alpha=0.05$) and effect size, respectively. In Table~\ref{tbl.stats} the three columns to the right show the $\hat{A}_{12}$ values for all algorithm combinations, where this value is significant and '-' otherwise. Thus, the higher the value in the $_{\Genetic}^{\Random}$ column, indicates that $\Random$ is slower in comparison with $\Genetic$. Depending on the actual $\hat{A}_{12}$ value, the effect size can be small (0.56), medium (0.64), and large (0.71).\\
\framebox{
  \parbox[t][0.65cm]{0.95\linewidth}{
    Both $\Genetic$ and $\RGenetic$ largely outperform $\Random$ in 50\% cases generally across the board and 67\% on the difficult functions.
  }
}

@@

\textbf{Threats to Validity:}
The \emph{construct validity} threat comes from the choice of measurements: branch coverage and execution time. Although, the branch coverage is a common criterion used in practice, the actual fault finding capability is a stronger indication of testing adequacy. Reaching a target branch is only part of the problem: we also need an oracle to discriminate the program's output. In this paper, we rely on so called ``natural'' oracles leading to exceptions. Using a stronger oracle for validation is future work. Another practical aspect of the test generation is the test case size, since it can both complicate the readability and increase the execution time. To mitigate this issue we configured the GA to start evolution with the candidates of small sizes, which helps to prevent data explosion. Additional measure that could be applied is a minimization procedure, aka \emph{hierarchical delta debugging}~\cite{misherghi2006hdd} and its variations~\cite{herfert2017automatically}, to the GA result.

### 2018-08-12 doc/issre2018/example.tex
Working on camera-ready. Fixed all comments by Doatse.

Let us consider the JS function \texttt{isGameFinished} in Figure~\ref{code.isGameFinished} taken from the web game \emph{sudoku}~\cite{sudoku}. This function checks if a given sudoku solution is valid. First, it locates a DOM element \texttt{obj} representing the game field (line~\ref{isGameFinished.getSudoku}). Then, it collects all children \emph{DIV} elements of \texttt{obj} (line~\ref{isGameFinished.getDivs}). Each \emph{DIV} corresponds to an input square of the game (line~\ref{isGameFinished.if1.begin}) and consists of two \emph{SPAN} tags. The first \emph{SPAN} contains the value entered by the player, whereas the second one (hidden by default) stores the expected (correct) value for that square. In the for-loop in lines~\ref{isGameFinished.inFor.begin}-\ref{isGameFinished.outFor}, we iterate through the whole collection of the input squares. Once, we encounter a square with two unequal values (line~\ref{isGameFinished.if2.begin}), we terminate the loop and call the game unfinished (line~\ref{isGameFinished.unfinished}), otherwise finished.

Suppose now that we would like to unit test this function. We aim to maximize the branch coverage as a common testing criteria~\cite{zhu1997software}. That is, for each individual branch of the function under test (FUT) \texttt{isGameFinished}, we need to construct a test input that covers the branch and leads to the normal termination of the FUT. A JS function can take \emph{explicit} input arguments (in our case none), but it can also accept \emph{implicit} arguments such as a DOM state. In general, in order to produce a complete test for our function we have to construct both an appropriate DOM and input arguments. The tests should be decoupled from the application's HTML. Moreover, such HTML could still be under active development. Thus, the developers have to complement test with DOM fixtures.

@@

As we have just witnessed, even for such a relatively simple example, the test data generation can be far from trivial procedure because it requires the deep understanding of the program semantics. In practice, web developers often have to deal with third-party code, including a large number of JS libraries freely available on the GitHub, which are commonly untested and poorly documented. At the same time, unit testing is the most fundamental level of testing and the easiest to implement. It saves the code from unexpected regressions and often documents the desired behavior of a program. Thus, in industry unit testing has became a standard practice in web development.

>>>>

Let us consider the JS function \texttt{isGameFinished} in Figure~\ref{code.isGameFinished} taken from the web game \emph{sudoku}~\cite{sudoku}. This function checks if a candidate sudoku solution is valid. First, it locates a DOM element \texttt{obj} representing the game field (line~\ref{isGameFinished.getSudoku}). Then, it collects all children \emph{DIV} elements of \texttt{obj} (line~\ref{isGameFinished.getDivs}). Each \emph{DIV} corresponds to an input square of the game (line~\ref{isGameFinished.if1.begin}) and consists of two \emph{SPAN} tags. The first \emph{SPAN} contains the value entered by the player, whereas the second one (hidden by default) stores the expected (correct) value for that square. In the for-loop in lines~\ref{isGameFinished.inFor.begin}-\ref{isGameFinished.outFor}, we iterate through the whole collection of the input squares. Once, we encounter a square with two unequal values (line~\ref{isGameFinished.if2.begin}), we terminate the loop and call the game unfinished (line~\ref{isGameFinished.unfinished}), otherwise finished.

Suppose now that we would like to unit test this function. We aim to maximize the branch coverage as a common testing criterion~\cite{zhu1997software}. That is, for each individual branch of the function under test (FUT) \texttt{isGameFinished}, we need to construct a test input that covers the branch and leads to the normal termination of the FUT. A JS function can take \emph{explicit} input arguments (in our case none), but it can also accept \emph{implicit} arguments such as a DOM state. In general, in order to produce a complete test for our function we have to construct both an appropriate DOM and input arguments. The tests should be decoupled from the application's HTML. Moreover, such HTML could still be under active development. Thus, the developers have to complement test with DOM fixtures.

@@

As we have just witnessed, even for such a relatively simple example, the test data generation can be far from trivial procedure because it requires the deep understanding of the program semantics. In practice, web developers often have to deal with third-party code, including a large number of JS libraries freely available on GitHub, which are commonly untested and poorly documented. At the same time, unit testing is the most fundamental level of testing and the easiest to implement. It saves the code from unexpected regressions and often documents the desired behavior of a program. Thus, in industry unit testing has become a standard practice in web development.
